### 2019-09-04 paper draft.Rmd

---
title: "FATE summary results"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(Rcpp)
library(ewidata)
library(knitr)
library(reshape2)
library(rstan)
library(tibble)
library(dplyr)
library(gtools)
library(bayesdfa)
library(assertthat)
library(rstanarm)

```

## Results

This is the initial write-up on results for the FATE paper.

### Climate

The best model for GOA climate was the two-trend, independent error model. Trend one positively loaded SST and the Papa advection index, and negatively loaded GAK1 20m salinity. Trend two negatively loaded the upwelling indices and positively loaded the SLP gradient (Fig. 1A).

#### Fig. 1A
GOA climate trend loadings.

```{r, echo=F}

GOA.clim <- readRDS("GOA_clim_2_trends_original_data.rds") # read in GOA climate model

# recover the years and names for plots!
goa.clim <- read.csv("updated goa climate data.csv")

sub_data <- goa.clim %>%
  select(-ssh, -wind.stress) %>%
  gather(key="code", value="value", -year)

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.clim$best_model)

# set new names
new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")

# bespoke plotting code
loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
loadings.2 <- as.data.frame(rotated$Z_rot[,,2])
names(loadings.1) <- names(loadings.2) <- names

# drop columns that don't meet threshold
thresh <- 0.9

f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))

prop1 <- apply(loadings.1, 2, f)
prop2 <- apply(loadings.2, 2, f)

# so! only Trend 1 includes loadings > 0.9! 
# this justifies fitting/retaining only a 1-trend model!

# load the 1-trend object
GOA.clim <- readRDS("GOA_clim_1_trend_original_data.rds") # read in GOA climate model

# recover the years and names for plots!
goa.clim <- read.csv("updated goa climate data.csv")

sub_data <- goa.clim %>%
  select(-ssh, -wind.stress) %>%
  gather(key="code", value="value", -year)

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.clim$best_model)

# set new names
new.names <- c("East.spring.SST", "East.winter.SST", "SLP.gradient", "Papa.advection", "GAK1.salinity", "West.spring.SST", "West.winter.SST", "Downwell.54.134", "Downwell.57.137", "Downwell.60.146", "Downwell.60.149")

pdf("goa.climate.loadings.pdf", 6,6)
print(plot_loadings(rotated, names = new.names, threshold=0.8))
dev.off()

plot.trend <- data.frame(year=1950:2019, trend=as.vector(rotated$trends_mean),
                         lo=as.vector(rotated$trends_lower),
                         hi=as.vector(rotated$trends_upper))

plot.b <- ggplot(plot.trend, aes(year, trend)) +
  theme_bw() +
  geom_line(color=cb[3]) +
  geom_ribbon(aes(ymin=lo, ymax=hi), fill=cb[3], alpha=0.4) +
  geom_hline(yintercept = 0) + ylab("Trend value") +
  theme(axis.title.x = element_blank()) + 
  ggtitle("b) Trend")

png("combined GOA climate loading and trend plot.png", 6, 3, units = "in", res=300)

ggpubr::ggarrange(plot.a, plot.b, ncol=2)

dev.off()

# back to the bespoke code!
loadings.1 <- as.data.frame(rotated$Z_rot[,,1])
names(loadings.1)  <- new.names

# drop columns that don't meet threshold
# first, figure out the proportion above/below 0
f <- function(x) max(sum(x<0)/length(x), sum(x>0)/length(x))

prop1 <- apply(loadings.1, 2, f)
prop1
keep <- prop1 > 0.95

loadings.1 <- loadings.1[,keep]

# plot
plot.load <- loadings.1 %>%
  gather()

rank <- tapply(plot.load$value, plot.load$key, mean)
plot.load$rank <- rank[match(plot.load$key, names(rank))]

plot.load$key <- reorder(plot.load$key, plot.load$rank)

# set pallette
cb <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

plot.a <- ggplot(plot.load, aes(key, value)) +
  theme_bw() +
  geom_violin(fill=cb[3]) +
  geom_hline(yintercept = 0) +
  theme(axis.title.y = element_blank()) + ylab("Loading") +
  coord_flip() +
  ggtitle("a) Loadings")

modelfit <- GOA.clim$best_model

  n_ts <- dim(modelfit$data)[1]
  n_years <- dim(modelfit$data)[2]
  pred <- predicted(modelfit)
   res.df <- data.frame(ID = rep(seq_len(n_ts),
                                 n_years),
                        Time = sort(rep(seq_len(n_years), n_ts)), 
                        mean = c(t(apply(pred, c(3, 4), mean))), 
                        lo = c(t(apply(pred, c(3, 4), quantile, 0.025))),
                        hi = c(t(apply(pred, c(3, 4), quantile, 0.975))), 
                        y = c(modelfit$data))

        res.df$ID <- new.names

  res.df$year <-  rep(1950:2019, each=n_ts)

  res.df$mean.res <- res.df$y - res.df$mean
  res.df$lo.res <- res.df$y - res.df$lo
  res.df$hi.res <- res.df$y - res.df$hi

 plot.these <- names(keep)[keep]

 res.df <- res.df %>%
   filter(ID %in% plot.these)

  ggplot(res.df, aes_string(x = "year", y = "mean.res")) + geom_ribbon(aes_string(ymin = "lo.res", 
        ymax = "hi.res"), fill=cb[3], alpha = 0.4) + geom_line(color=cb[3]) + facet_wrap("ID", 
        scales = "free_y") + xlab("Year") + ylab("") + geom_hline(yintercept = 0, col=cb[2]) +
        geom_vline(xintercept = c(1988.5, 2014), lty=2)

# acf plot
    index <- unique(res.df$ID)

    par(mfrow=c(2,4))
    for(i in 1:length(index)){
     # i <- 1
      temp <- res.df %>%
        filter(ID==index[i])
      temp <- na.omit(temp)
      acf(temp$mean.res, main=index[i]) 
    }

# so...no autocorrelation concerns with residuals

# look at rolling correlations between fitted and observed values, as in Ecology paper
# using 25-yr rolling windows!

    dfa.cor <- data.frame()

    for(i in 1:length(index)){    
      # i <- 1 
      temp <- res.df %>%
        filter(ID==index[i])

       for(ii in 13:58){
       #  ii <- 13
          win <- temp[(ii-12):(ii+12),]
          temp.cor <- data.frame(ID=unique(win$ID),
                                 year=win$year[13], 
                                 window=25,
                                 mean.cor=cor(win$y, win$mean, use="p"),
                                 n.cor=sum(!is.na(win$y)))
          dfa.cor <- rbind(dfa.cor, temp.cor)
       }  }

# now the 11-yr windows
     for(i in 1:length(index)){    
      # i <- 1 
      temp <- res.df %>%
        filter(ID==index[i])
             for(ii in 6:65){
       #  ii <- 13
          win <- temp[(ii-5):(ii+5),]
          temp.cor <- data.frame(ID=unique(win$ID),
                                 year=win$year[6],
                                 window=11,
                                 mean.cor=cor(win$y, win$mean, use="p"),
                                 n.cor=sum(!is.na(win$y)))
          dfa.cor <- rbind(dfa.cor, temp.cor)
             }  }

      # and the 15-yr windows
    for(i in 1:length(index)){    
      # i <- 1 
      temp <- res.df %>%
        filter(ID==index[i])

             for(ii in 8:63){
       #  ii <- 13
          win <- temp[(ii-7):(ii+7),]
          temp.cor <- data.frame(ID=unique(win$ID),
                                 year=win$year[8],
                                 window=15,
                                 mean.cor=cor(win$y, win$mean, use="p"),
                                 n.cor=sum(!is.na(win$y)))
          dfa.cor <- rbind(dfa.cor, temp.cor)
       }  }

# remove salinity correlations with < 10 obs
dfa.cor <- dfa.cor %>%
  filter(n.cor>=10)
 dfa.cor$window <- as.factor(dfa.cor$window)

ggplot(dfa.cor, aes(year, mean.cor, color=window)) + 
  theme_bw() + theme(axis.title.x = element_blank()) +
  geom_line() + facet_wrap(~ID, 
        scales = "free_y") + xlab("Year") + ylab("Correlation")

ggsave("goa climate dfa correlations 11 15 25 year windows.png", height = 4, width=6, units="in")

# and plot 15-yr windows
ggplot(filter(dfa.cor, window==15), aes(year, mean.cor)) + 
  theme_bw() + theme(axis.title.x = element_blank()) +
  geom_line(color=cb[3]) + facet_wrap(~ID, 
        scales = "free_y") + xlab("Year") + ylab("Correlation")

ggsave("goa climate dfa correlations 15 year windows.png", height = 4, width=6, units="in")

# and plot 25-yr windows
ggplot(filter(dfa.cor, window==25), aes(year, mean.cor)) + 
  theme_bw() + theme(axis.title.x = element_blank()) +
  geom_line(color=cb[3]) + facet_wrap(~ID, 
        scales = "free_y") + xlab("Year") + ylab("Correlation")

ggsave("goa climate dfa correlations 25 year windows.png", height = 4, width=6, units="in")

```

### Fig. 1B
GOA climate trends

```{r, echo=F}
# trends
# pdf("goa.climate.trends.pdf", 6,6)
# print(plot_trends(rotated, years = as.numeric(colnames(Y))))
# dev.off()
```

Find swans.

```{r, echo=F}
xx <- find_swans(rotated, threshold = 0.01, plot=F)
xx$year <- 1950:2019
xx <- filter(xx, below_threshold == T)
xx
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

# # repeat with 2nd and third trends, e.g.
# set.seed(99)
# f2 = find_regimes(y = y[2,], sds = sd_y[2,], max_regimes = 3)
# 
# # # repeat with 2nd and third trends, e.g.
# set.seed(99)
# f3 = find_regimes(y = y[3,], sds = sd_y[3,], max_regimes = 3)
```

Regime results for Trend 1. The 2-regime model is best-supported, no evidence at this point for a switch to a novel, warmer state.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

## GOA biology results.

One trend model was the best.

```{r, echo=F}
GOA.biol <- readRDS("GOA_biol_9.4.19.rds") 
GOA.biol$summary
```

### Fig. 2A

```{r, echo=F}

# reload the version of data used
GOA <- read.csv("updated GOA biology data.csv")

# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.biol$best_model)

new.names <- c("Chignik.coho", "Chignik.pink", "Chiniak.eulachon", "Chiniak.jellyfish", "Chiniak.shrimp", "Cook.Inlet.coho", "Kodiak.coho", "Kodiak.pink", "Pavlof.capelin", "Pavlof.eulachon", "Pavlof.jellyfish", "Pavlof.shrimp", "Prince.William.coho", "Southeast.coho", "Southeast.herring", "Southeast.pink", "South.Peninsula.coho", "South.Peninsula.pink")

# loadings
pdf("goa.biol.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.biol.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the GOA biology trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}
# first, define the function
trend_lm <- function(rotated_modelfit,
  all_years, selected_years,
  climate_dat = NULL,
  trend_number = 1, samples = 50,
  ITER = 1000, CHAINS = 1, n_eff = 50, rhat = 1.1, show_plot = TRUE) {

  total_mcmc <- dim(rotated_modelfit$trends)[1]
  draws <- sample(seq_len(total_mcmc), samples)

  yrs_i <- which(all_years %in% selected_years)
  climate_dat <- dplyr::filter(climate_dat, Year %in% selected_years)

  predictor_name = names(climate_dat)[which(names(climate_dat) != "Year")]
  formula_str = paste("dfa_trend_draw ~", predictor_name[1])
  if(length(predictor_name) > 1) {
    for(i in 2:length(predictor_name)) {formula_str = paste(formula_str, predictor_name[i], sep=" + ")}
  }
  formula <- as.formula(formula_str)

  post <- list()
  for (i in seq_along(draws)) {
    dd = climate_dat
    dd$dfa_trend_draw = rotated_modelfit$trends[draws[i], trend_number, yrs_i]
    m_temp <- stan_glm(formula, data = dd,
      iter = ITER, chains = CHAINS)
    assert_that(all(m_temp$stan_summary[,"Rhat"] < rhat))
    assert_that(all(m_temp$stan_summary[,"n_eff"] > n_eff))
    post[[i]] <- as.data.frame(as.matrix(m_temp))
  }
  post_all <- dplyr::bind_rows(post)

  if (show_plot) {
    #par(mfrow = c(1, 2))
    #plot(post_all[,predictor], type = "l", col = "#00000080", lwd = 0.5)
    #abline(v = seq(0, ITER * length(draws), ITER), col = "#FF000050")
    #plot(density(post_all[,predictor]))
    #abline(v = 0, col = "#00000050")
  }
  invisible(post_all)
}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
  selected_years = 1972:2016, climate_dat=cli_dat)

names(mod)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
summary(mod)

# get 95% CIs

for(i in 2:4){
x <- sort(mod[,i])
print(names(mod)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. 3A goa.biol.on.sst.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
dev.off()

```

Now the PDO.

```{r, echo=F}

pdo <- read.csv("pdo.csv")

win.pdo <- pdo$NDJFM

years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
  selected_years = 1972:2016, climate_dat=cli_dat)

names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")

summary(mod2)

# get 95% CIs

for(i in 2:4){
x <- sort(mod2[,i])
print(names(mod2)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. 3B goa.biol.on.pdo.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
dev.off()

```

And the NPGO.

```{r, echo=F}

# 
# npgo <- read.csv("winter.npgo.csv")
# 
# years <- as.numeric(colnames(GOA.biol$best_model$data)) # these are the years for the biology model - 1972:2016
# 
# # set up climate data
# cli_dat <- data.frame(Year=npgo$YEAR,NPGO=npgo$NDJFM)
# 
# cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(NPGO = arm::rescale(NPGO))
# 
# # cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))
# 
# cli_dat$slope <- 1
# cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
# cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
# cli_dat$slope <- as.factor(cli_dat$slope)
# 
# # rename rotated
# r <- rotated
# 
# Y = rnorm(length(cli_dat$slope))
# cli_dat = as.data.frame(model.matrix(Y ~ Year + NPGO:slope, data=cli_dat))
# # replace ':' in name for formula
# names(cli_dat) = gsub(":","",names(cli_dat))
# # remove named intercept
# cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]
# 
# mod2 = trend_lm(rotated_modelfit = r, all_years = 1972:2016,
#   selected_years = 1972:2016, climate_dat=cli_dat)
# 
# 
# summary(mod2)
# 
# names(mod2)[2:4] <- c("1972-1988", "1989-2013", "2014-2016")
# 
# melt.mod <- melt(mod2, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
# names(melt.mod)[3] <- "era"
# 
# pdf("Fig. 3C goa.biol.on.npgo.pdf", 6, 4)
# ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) NPGO")
# dev.off()

```

## GOA ichthyo results.

One trend model was again the best.

```{r, echo=F}
GOA.ich <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_ich_7.19.18.one.trend.rds") 
GOA.ich$summary
```

### Fig. _A

```{r, echo=F}

# reload the version of data used
# load updated ewidata!
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
unique(GOA$code)

# limit to ichthy

keep <- grep("ICH", GOA$code)

GOA <- GOA[keep,]

# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.ich$best_model)

new.names <- c("sandlance", "ronquils", "walleye.pollock", "Pacific.cod", "flathead.sole", "n.rock.sole", "rockfish")

# loadings
pdf("goa.ich.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.ich.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(11)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the ichthyo trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.ich$best_model$data))

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

# this time series is too short to divide up into eras!
cli_dat$slope <- 1
#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = years,
  selected_years = years, climate_dat=cli_dat)

summary(mod)

# get 95% CIs
x <- sort(mod[,2])
print(names(mod)[2])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))

# dropping this b/c we have no eras for this short TS!
# melt.mod <- melt(mod, measure.vars=c("1972-1988", "1989-2013", "2014-2016"))
# names(melt.mod)[3] <- "era"

pdf("Fig. x goa.ich.on.sst.pdf", 4, 5)
ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
dev.off()

```

Now the zoops model.

```{r, echo=F}
GOA.zoop <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/GOA_zoop_7.19.18.one.trend.rds") 
GOA.zoop$summary
```

```{r, echo=F}

# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# limit to GOA
GOA <- ewidata[ewidata$system %in% c("WGOA", "EGOA"),]
unique(GOA$code)

# drop salmon
drop <- grep("PI", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("CO", GOA$code)
GOA <- GOA[-drop,]

# drop climate
drop <- grep("AKCLIM", GOA$code)
GOA <- GOA[-drop,]

unique(GOA$code)

# drop Pavlof and Chiniak and herring and kittiwakes
drop <- grep("PAV", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("CHI", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("HER", GOA$code)
GOA <- GOA[-drop,]

drop <- grep("BLKI", GOA$code)
GOA <- GOA[-drop,]

# drop total icy zoops!
drop <- grep("ICY.ZOOP.TOTDEN", GOA$code)
GOA <- GOA[-drop,]

# and drop ichthyoplankton
drop <- grep("ICH", GOA$code)
GOA <- GOA[-drop,]

unique(GOA$code)
# recover the years and names for plots!
sub_data = GOA

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(GOA.zoop$best_model)

new.names <- c("CPR.copepods", "CPR.diatoms", "Icy.Str.amphipods", "Icy.Str.euphausids", "Icy.Str.gastropods", "CPR.mesozooplankton", "Seward.A.calanoid", "Seward.A.euphausid", "Seward.B.calanoid", "Seward.B.euphausid", "Seward.LA.euphausid", "Seward.LB.calanoid", "Seward.LB.euphausid", "Seward.calanoid.size", "Seward.euphausid.size", "Seward.LA.calanoid")

# loadings
pdf("goa.zoop.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()
```

```{r, echo=F}
# trends

pdf("goa.zoop.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Test for regime probability. Use n=3 maximum states.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(11)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)

```

Next step is to examine relationships between SST and the zoop trend.

We're interested in testing the hypothesis that the slope on sst was greater during 1972:1988 (high Aleutian Low variance) than during 1989:2013 (low Aleutian Low variance). Also interested in asking which of these periods the 2014-106 heat wave years are most similar to.

```{r, include=F}

#  load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_GOA", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, mean)

years <- as.numeric(colnames(GOA.zoop$best_model$data))

# set up climate data
cli_dat <- data.frame(Year=as.numeric(names(win.sst)),SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
#cli_dat$slope[cli_dat$Year %in% 1997:2013] <- 2
#cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod = trend_lm(rotated_modelfit = r, all_years = years,
  selected_years = years, climate_dat=cli_dat)

summary(mod)

# get 95% CIs
x <- sort(mod[,2])
print(names(mod)[2])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))

pdf("Fig. x goa.zoop.on.sst.pdf", 4, 5)
ggplot(mod, aes(x=SST)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + xlab("Slope")
dev.off()

```

# Now the EBS results.

```{r, echo=F}

EBS.clim <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_clim_6.19.18.rds") # read in EBS climate model

EBS.clim$summary
```

And the trends and loading plots.

```{r, echo=F}

# recover the years and names for plots!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_EBS", ewidata$code)

ewidata <- ewidata[keep,]
sub_data = ewidata

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(EBS.clim$best_model)

new.names <- c("Bottom.temp", "Spring.SST", "Winter.SST", "Summer.NW.wind", "Summer.SE.wind", "Winter.NW.wind", "Winter.SE.wind", "Ice.57.75N", "Ice.58.75N", "Ice.64.25N")

# loadings
pdf("ebs.clim.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()

# trends
pdf("ebs.clim.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```

Find swans.

```{r, echo=F}
xx <- find_swans(rotated, threshold = 0.01, plot=F)
xx$year <- 1950:2016
xx <- filter(xx, below_threshold == T)
xx
```

Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

```

Regime results for Trend 1. The 2-regime model is best-supported.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

## EBS biology results.

```{r, echo=F}
EBS.biol <- readRDS("/Users/MikeLitzow/Documents/R/Fate/fate-ewi/EBS_biol_6.19.18.rds") 
EBS.biol$summary
```

Now plot trends and loadings.

```{r, echo=F}

# reload the version of data used
# load updated ewidata!
ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))

# first, EBS biology
codes <- unique(ewidata$code)

# limit to EBS
EBS <- ewidata[ewidata$system == "EBS",]

# and start in 1975, when data beyond salmon are first available
EBS <- EBS[EBS$year>=1975,]

unique(EBS$code)

# drop clim
drop <- grep("AKCLIM", EBS$code)
EBS <- EBS[-drop,]

# drop depth
drop <- grep("DEPTH", EBS$code)
EBS <- EBS[-drop,]

# recover the years and names for plots!
sub_data = EBS

melted = melt(sub_data[, c("code", "year", "value")], id.vars = c("code", "year"))
Y <- dcast(melted, code ~ year)
names = Y$code
Y = as.matrix(Y[,-which(names(Y) == "code")])

rotated = rotate_trends(EBS.biol$best_model)

new.names <- c("Alaska.place.lat", "Bristol.Bay.coho", "Bering.flounder.lat", "St.George.kitt.phen", "St.Paul.kitt.phen", "St.Paul.kitt.product", "com.eider.eggs/nest", "com.eider.hatch.date", "com.eider.success", "St.George.com.murre.phenol", "St.Paul.com.murre.phenol", "St.George.com.murre.product", "St.Paul.com.murre.product", "juv.cod.cpue", "juv.pollock.cpue", "flathead.sole.lat", "gl.gull.eggs/nest", "gl.gull.hatch.date", "gl.gull.success", "Grnlnd.turbot.lat", "Pac.cod.lat", "Pac.halibut.lat", "red.king.crab.lat", "rock.sole.lat", "shortf.eelpout.lat", "snow.crab.lat", "spec.eider.eggs/nest", "spec.eider.hatch.date", "spec.eider.success", "sturgeon.poacher.lat", "Tanner.crab.lat", "St.George.tb.murre.phenol", "St.Paul.tb.murre.phenol", "St.George.tb.murre.product", "St.Paul.tb.murre.product", "walleye.pollock.lat", "wattled.eelpout.lat", "yellowfin.sole.lat")

# loadings
pdf("ebs.biol.loadings.pdf", 6, 6)
print(plot_loadings(rotated, names = new.names, threshold=0.9))
dev.off()

```

```{r, echo=F}
# trends

pdf("ebs.biol.trend.pdf", 6,6)
print(plot_trends(rotated, years = as.numeric(colnames(Y))))
dev.off()
```
Test for regime probability. Use n=3 maximum states to rule out a switch to a warmer state.

```{r, include=F}

y = apply(rotated$trends, c(2,3), mean)
sd_y = apply(rotated$trends, c(2,3), sd)

# run each trend sequentially through the regimes
# code to identify (1) number of changes and (2)
# where those change points occur.
set.seed(99)
f1 = find_regimes(y = y[1,], sds = sd_y[1,], max_regimes = 3)

```

Regime results for Trend 1. The 2-regime model is best-supported.

```{r, echo=F}
print(f1$table) # this shows 2-regime model is best
plot_regime_model(f1$best_model)
```

Next step is to examine relationships between SST and the EBS biology trend.

```{r, message=FALSE, results='hide', echo=F}

# load sst as the explanatory variable.

ewidata <- as.tibble(read.csv("ewidata.csv", row.names=1))
keep <- grep("AKCLIM_EBS", ewidata$code)
clim.var <- ewidata[keep,]  
keep <- grep("win.sst", clim.var$code)
clim.var <- clim.var[keep,]

win.sst <- tapply(clim.var$value, clim.var$year, identity)

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016

cli_dat <- data.frame(Year=as.numeric(names(win.sst)), SST=win.sst)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(SST = arm::rescale(SST))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + SST:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(4)
mod = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")
summary(mod)

# get 95% CIs

for(i in 2:4){
x <- sort(mod[,i])
print(names(mod)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.sst.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="A) SST")
dev.off()
```

Now the PDO.

```{r, message=FALSE, results='hide', echo=F}

pdo <- read.csv("pdo.csv")

win.pdo <- pdo$NDJFM

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1972:2016

# set up climate data
cli_dat <- data.frame(Year=pdo$YEAR,PDO=win.pdo)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(PDO = arm::rescale(PDO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + PDO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod2 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod2)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")

summary(mod2)

# get 95% CIs

for(i in 2:4){
x <- sort(mod2[,i])
print(names(mod2)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

melt.mod <- melt(mod2, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.pdo.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) PDO")
dev.off()

```

And the winter AO.

```{r, echo=F}

ao <- read.csv("winter.ao.csv")

win.ao <- ao$NDJFM

years <- as.numeric(colnames(EBS.biol$best_model$data)) # these are the years for the biology model - 1975:2016

# set up climate data
cli_dat <- data.frame(Year=ao$YEAR,AO=win.ao)

cli_dat <- filter(cli_dat, Year %in% years) %>% mutate(AO = arm::rescale(AO))

# cli_dat$regime = as.factor(sample(c(0,1,2), size=20, replace=T))

cli_dat$slope <- 1
cli_dat$slope[cli_dat$Year %in% 1989:2013] <- 2
cli_dat$slope[cli_dat$Year %in% 2014:2016] <- 3
cli_dat$slope <- as.factor(cli_dat$slope)

# rename rotated
r <- rotated

Y = rnorm(length(cli_dat$slope))
cli_dat = as.data.frame(model.matrix(Y ~ Year + AO:slope, data=cli_dat))
# replace ':' in name for formula
names(cli_dat) = gsub(":","",names(cli_dat))
# remove named intercept
cli_dat = cli_dat[,-which(names(cli_dat)=="(Intercept)")]

set.seed(99)
mod3 = trend_lm(rotated_modelfit = r, all_years = 1975:2016,
  selected_years = 1975:2016, climate_dat=cli_dat)

names(mod3)[2:4] <- c("1975-1988", "1989-2013", "2014-2016")

summary(mod3)

# get 95% CIs

for(i in 2:4){
x <- sort(mod3[,i])
print(names(mod3)[i])
print(paste("median: ", median(x)))
print(paste("LCI: ", x[(0.025*25000)]))
print(paste("UCI: ", x[(0.975*25000)]))
}

# and proportion > 0
for(i in 2:4){
x <- sort(mod3[,i])
print(names(mod3)[i])
print(paste("prop > 0: ", sum(x > 0)/length(x)))
}

melt.mod <- melt(mod3, measure.vars=c("1975-1988", "1989-2013", "2014-2016"))
names(melt.mod)[3] <- "era"

pdf("Fig. x ebs.biol.on.ao.pdf", 6, 4)
ggplot(melt.mod, aes(x=value, color=era)) + geom_density() + geom_hline(yintercept=0) + geom_vline(xintercept = 0, lty=2) + labs(title="B) AO")
dev.off()
```

```{r}
head(melt.mod)
restricted <- filter(melt.mod, era %in% c("1975-1988", "1989-2013"))
restricted <- melt.mod
png("ebs community response to ao.png", 6,4, units="in", res=300)
ggplot(data=restricted, aes(fill=era, x=era, y=value)) + geom_violin() + coord_flip() + ggtitle("Community response to Arctic Oscillation") +
  geom_hline(yintercept = 0) + xlab("Era") + ylab("Regression coefficient") + theme_gray() + theme(legend.position = c(0.8, 0.5), legend.title = element_blank())
dev.off()
```